Supplement to: Optimal management of a stochastically varying population when policy adjustment is costly

### 2014-12-24 manuscripts/supplement.Rmd
minor edits from Paul, flushing out supplement

---
title: Optimal management of a stochastically varying population when policy adjustment is costly
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTKEcon
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul Armsworth 
    affiliation: UTKEEB

@@

\appendix

@@

Though this results in an effect that is less pronounced, the stock dynamics clearly reflect the same general trends observed in Figure 3: the quadratic costs ($\Pi_2$) show a clear smoothing of the variation in stock dynamics, while the fixed fee $\Pi_3$ results in exaggerated fluctionations in the stock sizes.  It is worth noting how changes in policy costs alone, such as introducing or rising fixed costs, can have important knock-on effects to population dynamics such as increased variance, a recent phenomenon observed in important fisheries that has recieved much attention elsewhere [@Anderson2008].

```{r StockDynamics, dependson="tidy", fig.cap="Example realization of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=5, fig.width=8, cache = fig.cache, dev=c("pdf", "svg", "png")}

@@

```{r StockStats, fig.cap="Variance and autocorrelation of stock dynamics, as a function of the penalty coefficient.  Adjustment penalty is calculated as percentage of the maximum expected $NPV$ in the basic case with no policy adjustment costs.", fig.height=8, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png")}

@@

@@

Figure 3 of the main text shows a single stochastic realization.  Here we provide a sample of further realizations (including both stock and harvest dynamics).  Full

```{r Figure_S2, dependson="tidy", fig.cap="Multiple realizations of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=10, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png"), cache=FALSE}

@@

write.csv(ex, "components/data/figure_s2.csv")
```

@@

```{r quadraticcosts}

@@

```{r Figure_S1, dev=c('pdf', 'svg'), cache=FALSE}

@@

Like the $\Pi_2$ adjustment cost based penalty, this quadratic cost based penalty shows periods of contant, strictly positive policy only `r ave * 100`% of the time, and exhibits the same kind of smoothing of more small adjustments that is seen in $\Pi_2$ scenario.

@@

Figure S3 extends Figure 5 of the main text across all three functional forms of adjustment costs, $\Pi_{1-3}$. To reduce overplotting of Figure 5 we have also separated the distributions of dockside revenue and adjustment fees by scenario, such that the top row shows the distributions when the policy accounts for adjustment costs, and the bottom shows the policy when it does not (that is, the original @Reed1979 model applied to the scenario of costly adjustment).

```{r Figure_S3, fig.cap="Distribution of profits from fishing and costs paid to adjustment.  Distributions are shown when managing a system in which the adjustment costs present are determined by the model indicated, and compared to the same costs and profits of managing the same system using a Reed model, which ignores adjustment costs, and thus tends to pay much higher fees for adjustment.  In contrast, the difference in profit after accounting for costs is only marginally less than the value derived by managing the population to optimize profits when ignoring adjustment costs (Reed model).", fig.width=8, fig.height=6, dependson="histogram-calc", dev=c("pdf", "svg", "png"), cache=FALSE}

@@

write.csv(hist_dat, "components/data/figure5.csv")
```

Figure S4 extends the results shown in Figure 6 across different severities of adjustment cost, totalling from 10% up to 25% of the NPV of the stock. While the same qualitative pattern is observed in all cases, it is strongest in the higher penalties.

```{r Figure_S4, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue), shown by penalty function over increasing magnitudes of cost.  All values are relative to the cost-free adjustment optimum, NPV0.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. The severity of the latter mistake increases sharply as adjustment costs make a larger fraction of the total costs.", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}

ggplot(table1_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, ncol=2) + 
  scale_x_discrete(labels = c('L1' = substitute(paste(Pi[1])),
                              'L2' = substitute(paste(Pi[2])),
                              'fixed' = substitute(paste(Pi[3])))) + 
  xlab("Penalty function")
write.csv(table1_long, "components/data/figureS4.csv")
```

@@

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure S5 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10% (panel A)
less than optimal free based on the $c_i$ calibration. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

@@

```{r Figure_S5, fig.cap="Mismatches. X axis labels indicate the model used to derive the policy, followed by the model used to simulate the penalty. Blue bars show the net present value of that scenario.  Compare to red bars, which show the consequence of using a policy that ignores costs all together in the same scenario (i.e. costs as indicated by second code), rather than assuming the policy indicated by the first code in the pair.  All values are scaled to a fraction of the adjustment free net present value.", dev=c("pdf", "svg", "png"), fig.width=7}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

ggplot(filter(table2_long, reduction==0.25), 
  aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
write.csv(table2_long, "components/data/figureS5.csv")
```

These patterns hold across the different size reductions shown (from 10%
to 25%, Figure S6 panels a-d respectively), though net present value becomes negative under the quadratic
form for the larger penalties as costs paid for adjustments exceed
profits derived from harvest.  (Though optimal solutions will always
avoid negative net present values, these scenarios in which policies are
applied under conditions other than assumed in the optimality calculation
have no such guarantee.) For each scenario, results are averaged over
100 replicates.

```{r Figure_S6, fig.cap="Effect of varying severity of adjustment costs on mismatches. Successive panels show increasing reduction of the net present value by increasing penalty coefficients: e.g. a policy in the first panel would lose only 10% (panel A) of its net present value to adjustment costs if it had the correct form of those costs, as determined by the choice of penalty coefficient.  Plots structured as in figure 7, which corresponds to the fourth panel", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}
ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
```

>>>>

---
title: "Supplement to: Optimal management of a stochastically varying population when policy adjustment is costly"
author: 
  - name: Carl Boettiger
    affiliation: cstar
    email: cboettig (at) gmail.com
    footnote: Corresponding author
  - name: Michael Bode
    affiliation: melbourne
  - name: James Sanchirico
    affiliation: UCD
  - name: Jacob LaRiviere
    affiliation: UTKEcon
  - name: Alan Hastings
    affiliation: UCD
  - name: Paul Armsworth 
    affiliation: UTKEEB

@@

```{r configure, include=FALSE, cache=FALSE}
library("methods")
library("knitr")
basename <- "supplement"
opts_chunk$set(fig.path = paste("components/figure/", basename, "-", sep=""),
               cache.path = paste("components/cache/", basename, "/", sep=""))
opts_chunk$set(cache = 2)
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, 
               comment = NA, verbose = TRUE, echo=FALSE)
# PDF-based figures
opts_chunk$set(dev='pdf')
fig.cache <- TRUE

library("rmarkdown")
library("pdgControl")
library("reshape2")
library("plyr")
library("ggplot2")
library("data.table")
library("pander")
library("cboettigR")
library("ggthemes")
library("snowfall")
library("tidyr")
library("dplyr")
options(digits=2)

theme_set(theme_tufte(base_size=12) + theme(axis.line = element_line()))
```

\appendix
\renewcommand{\thefigure}{S\arabic{figure}}

@@

Though this results in an effect that is less pronounced, the stock dynamics of Figure S1 reflect the same general trends observed in Main text Figure 3; e.g. the fixed fee $\Pi_3$ results in the most exaggerated fluctuations in the stock sizes.  It is worth noting how changes in policy costs alone, such as introducing or rising fixed costs, can have important knock-on effects to population dynamics such as increased variance, a recent phenomenon observed in important fisheries that has received much attention elsewhere [@Anderson2008].

```{r Figure_S1, dependson="tidy", fig.cap="Example realization of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=5, fig.width=8, cache = fig.cache, dev=c("pdf", "svg", "png")}

@@

```{r Figure_S2, fig.cap="Variance and autocorrelation of stock dynamics, as a function of the penalty coefficient.  Adjustment penalty is calculated as percentage of the maximum expected $NPV$ in the basic case with no policy adjustment costs.", fig.height=4, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png")}

@@

Figure S2 shows the same summary statistics illustrating the trends
of increasing adjustment costs on the _stock_ dynamics, instead of the
_harvest_ dynamics shown in Figure 4 of the main text.  Note that 
the variance in stock sizes under the quadratic penalty ($\Pi_2$) increases,
while in Figure 4 of the main text the corresponding variance in the
harvest decreases. Under costly adjustment, the undershooting behavior
of the $\Pi_2$ harvest policy cannot absorb as much of the natural 
fluctuation, resulting in increased variation in the stock sizes.

@@

Figure 3 of the main text shows only a single stochastic realization.  Here in
Figure S3 we provide a sample of further realizations (including both
stock and harvest dynamics).

```{r Figure_S3, dependson="tidy", fig.cap="Multiple realizations of optimal harvesting strategy and resulting fish stock sizes under the different functional forms of adjustment costs.", fig.height=10, fig.width=8, cache=fig.cache, dev=c("pdf", "svg", "png"), cache=FALSE}

@@

write.csv(ex, "components/data/figure_s3.csv")
```

@@

Figure S4 shows that adding quadratic terms to the cost of fishing effort directly, as in @Lewis1981 or @McGough2009, results in a similar smoothing effect as we observe in quadratic adjustment costs ($\Pi_2$), which contrasts sharply with the patterns we observe in other functional forms.  This case corresponds to a profit function defined as:

$$\Pi_4(N_t,h_t) = p h_t -  c_0 E_t - c_4 E_t^2$$

Where $c_4$ is calibrated as in the case of the other coefficients to induce the fixed reduction of net present value relative to $NPV_0$.

```{r quadraticcosts, include=FALSE, message=FALSE}

@@

```{r Figure_S4, dev=c('pdf', 'svg'), cache=FALSE, fig.cap="Quadratic costs to fishing effort show a similar pattern of smoothing to that which we observe in the quadratic adjustment cost model."}

@@

Like the $\Pi_2$ adjustment cost based penalty, this quadratic cost based penalty shows periods of constant, strictly positive policy only `r ave * 100`% of the time, and exhibits the same kind of smoothing of more small adjustments that is seen in $\Pi_2$ scenario.

@@

Figure S5 extends Figure 5 of the main text across all three functional forms of adjustment costs, $\Pi_{1-3}$. To reduce over-plotting of Figure 5 we have also separated the distributions of dockside revenue and adjustment fees by scenario, such that the top row shows the distributions when the policy accounts for adjustment costs, and the bottom shows the policy when it does not (that is, the original @Reed1979 model applied to the scenario of costly adjustment).

```{r Figure_S5, fig.cap="Distribution of profits from fishing and costs paid to adjustment.  Distributions are shown when managing a system in which the adjustment costs present are determined by the model indicated, and compared to the same costs and profits of managing the same system using a Reed model, which ignores adjustment costs, and thus tends to pay much higher fees for adjustment.  In contrast, the difference in profit after accounting for costs is only marginally less than the value derived by managing the population to optimize profits when ignoring adjustment costs (Reed model).", fig.width=8, fig.height=6, dependson="histogram-calc", dev=c("pdf", "svg", "png"), cache=FALSE}

@@

write.csv(hist_dat, "components/data/figureS5.csv")
```

Figure S6 extends the results shown in Figure 6 across different severities of adjustment cost, totaling from 10% up to 25% of the $NPV$ of the stock. While the same qualitative pattern is observed in all cases, it is strongest in the higher penalties.

```{r Figure_S6, fig.cap="Relative cost of ignoring adjustment costs when they are present ('ignoring', red) vs assuming the adjustment costs when they are absent ('assuming', blue), shown by penalty function over increasing magnitudes of cost.  All values are relative to the cost-free adjustment optimum, NPV0.  In each case, it is less costly to assume adjustment costs are present when they are not than it is to ignore them when they are present. The severity of the latter mistake increases sharply as adjustment costs make a larger fraction of the total costs.", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}

ggplot(table1_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, ncol=2) + 
  scale_x_discrete(labels = c('L1' = substitute(paste(Pi[1])),
                              'L2' = substitute(paste(Pi[2])),
                              'fixed' = substitute(paste(Pi[3])))) + 
  xlab("Penalty function")
write.csv(table1_long, "components/data/figureS6.csv")
```

@@

When policy costs are present, assuming the wrong functional form may or
may not be worse than ignoring the presence of policy costs all together.
Figure S7 considers several such cases.  The X axis labels indicate
first which policy is being used to drive the harvest decisions, and
then which penalty is actually being applied.  Thus $\Pi_1\_\Pi_2$ indicates
a policy calculated under the linear ($\Pi_1$) costs, when the reality
involves quadratic ($\Pi_2$) costs.  Though the coefficients of the ($\Pi_2$)
costs have been scaled such that the net effect is the same, the inferred
policies differ.  As a result, the value is slightly more than half of
the optimal free solution when the expected reduction should only be 10% (panel A)
less than optimal free based on the $c_i$ calibration. In this case, ignoring the costs all together
is still worse than using $\Pi_1$ costs -- the red bar is lower than the blue.

@@

```{r Figure_S7, fig.cap="Mismatches. X axis labels indicate the model used to derive the policy, followed by the model used to simulate the penalty. Blue bars show the net present value of that scenario.  Compare to red bars, which show the consequence of using a policy that ignores costs all together in the same scenario (i.e. costs as indicated by second code), rather than assuming the policy indicated by the first code in the pair.  All values are scaled to a fraction of the adjustment free net present value.", dev=c("pdf", "svg", "png"), fig.width=7}
who <- c("penalty_fn", "ignore_fraction", "mismatched_fraction", "reduction")
table2 <- arrange(mismatches_df[who], reduction) 
names(table2) = c("penalty.fn", "ignoring", "mismatched", "reduction")
table2_long <- melt(table2, id = c('penalty.fn', 'reduction'))
table2_long$reduction <- as.factor(table2_long$reduction)

ggplot(filter(table2_long, reduction==0.25), 
  aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
write.csv(table2_long, "components/data/figureS7.csv")
```

These patterns hold across the different size reductions shown (from 10%
to 25%, Figure S8 panels a-d respectively), though net present value becomes negative under the quadratic
form for the larger penalties as costs paid for adjustments exceed
profits derived from harvest.  (Though optimal solutions will always
avoid negative net present values, these scenarios in which policies are
applied under conditions other than assumed in the optimality calculation
have no such guarantee.) For each scenario, results are averaged over
100 replicates.

```{r Figure_S8, fig.cap="Effect of varying severity of adjustment costs on mismatches. Successive panels show increasing reduction of the net present value by increasing penalty coefficients: e.g. a policy in the first panel would lose only 10% (panel A) of its net present value to adjustment costs if it had the correct form of those costs, as determined by the choice of penalty coefficient.  Plots structured as in figure 7, which corresponds to the fourth panel", dev=c("pdf", "svg", "png"), fig.width=7, cache=FALSE}
ggplot(table2_long, aes(penalty.fn, value, fill = variable)) + 
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~reduction, scales="free_y") +
  scale_x_discrete(labels = c('L1_L2' = substitute(paste(Pi[1],"_", Pi[2])), 
                              'L2_L1' =  substitute(paste(Pi[2],"_", Pi[1])),
                              'L1_fixed' = substitute(paste(Pi[1],"_", Pi[3])), 
                              'fixed_L1' =  substitute(paste(Pi[3],"_", Pi[1])),
                              'fixed_L2' = substitute(paste(Pi[3],"_", Pi[2])), 
                              'L2_fixed' =  substitute(paste(Pi[2],"_", Pi[3])))) + 
  xlab("Penalty function")
```
